---
title: "Budweiser Case Study"
author: "Ravi Sivaraman"
date: "7/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Craft Beers and Breweries
An analysis of 2410 US craft beers and 558 breweries across United States. 
Budweiser CFO presented  dataset of 2410 US craft beers and 558 breweries  to us to analyze and find interesting data report that may be intersting prospect for Budweiser. 
We did an analysis of the beers and breweries dataset, and found lot of useful information that is benefitial to Budweiser. 


```{r include=FALSE}
library(dplyr)
library(ggplot2)
library(readr)
library(knitr)
library(kableExtra)
library(tidyr)
```

## Import Data

The section imports the data into R

```{r  echo=TRUE}
breweries <- read.csv("Breweries.csv")
beers <- read.csv("Beers.csv")
```

## 1 Breweries by State

```{r  echo=TRUE}
#-------------------------------------------------------------------------
#The following section of the code loads the breweries and selects only
#the State column.
#group by, groups the breweries by state
#summarize, adds the count of breweries by state
#brewbystate contains a dataframe that has State and number of breweries in that state
#-------------------------------------------------------------------------

brewbystate <- breweries %>% select(State) %>% group_by(State) %>% summarize(Count=n())

brewbystate <- brewbystate[order(brewbystate$Count, decreasing = TRUE),]
brewbystate$State = factor(brewbystate$State, levels = brewbystate$State[order(brewbystate$Count, decreasing = TRUE)])
```

The following list containst the list of breweries in United States by State.

```{r  echo=TRUE}
#This is to generate the table
kable(brewbystate) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"), full_width = F, position="left")
```

```{r  echo=TRUE}
#This is to plot
brewbystate %>%ggplot(aes(x=State,y=Count)) + geom_col(color="darkorange", fill="orange") + xlab("State") + ylab("Number of Breweries") + ggtitle("Count of Breweries By State") + theme(axis.text.x = element_text(angle = 90)) + geom_text(aes(label=Count, color="yellow"), vjust=-.5, size=3)
```

## 2 Merge Breweries and Brews data
```{r  echo=TRUE}
colnames(breweries)[1] = "Brewery_id"
brewandbeers = merge(breweries, beers, by ="Brewery_id")
colnames(brewandbeers)[1] = "ID"
colnames(brewandbeers)[2] = "Brewery"
colnames(brewandbeers)[5] = "Beer"
kable(brewandbeers[1:6,]) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"), full_width = F, position="left")
kable(tail(brewandbeers, keepnums = FALSE)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"), full_width = F, position="left")
```

## 3 Missing Values
There were 2,410 records missing values. 
1,005 were missing either IBU or ABV values so we populated these with min value as their respective median.
We are interested in median, so the value doesn't affect the median very much.

```{r  echo=TRUE}
#Replace NA with minimum value of ABV and IBU respectively.
minABV = min(brewandbeers$ABV, na.rm = TRUE)
brewandbeers <- brewandbeers %>% mutate(ABV=replace_na(ABV,minABV))

minIBU = min(brewandbeers$IBU, na.rm = TRUE)
brewandbeers <- brewandbeers %>% mutate(IBU=replace_na(IBU,minIBU))

```

## 4 Median ABV and IBU by State
The following list shows the median IBU and ABV by State

```{r  echo=TRUE}
# Computes the median ABV and median IBU
#Groups by State
medianstate<- brewandbeers %>% group_by(State) %>% summarize("Median ABV"=median(ABV), "Median IBU"=median(IBU))

kable(medianstate) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"), full_width = F, position="left")

#Plot the bar graph
colnames(medianstate)[2] ="MedianABV"
medianstateABV <- medianstate[order(medianstate$MedianABV, decreasing = TRUE),]
medianstateABV$State <- factor(medianstateABV$State, levels = medianstateABV$State[order(medianstateABV$MedianABV, decreasing = TRUE)])


medianstateABV %>%ggplot(aes(x=State,y=MedianABV)) + geom_col(color="darkorange", fill="orange") + xlab("State") + ylab("Median ABV") + ggtitle("Median ABV by State") + theme(axis.text.x = element_text(angle = 90)) + geom_text(aes(label=MedianABV), hjust=-.4, size=2.2) +   coord_flip()

```

Median IBU by State
```{r  echo=TRUE}

colnames(medianstate)[2] ="MedianIBU"
medianstateIBU <- medianstate[order(medianstate$MedianIBU, decreasing = TRUE),]
medianstateIBU$State <- factor(medianstateIBU$State, levels = medianstateIBU$State[order(medianstateIBU$MedianIBU, decreasing = TRUE)])


medianstateIBU %>%ggplot(aes(x=State,y=MedianIBU)) + geom_col(color="darkorange", fill="orange") + xlab("State") + ylab("Median IBU") + ggtitle("Median IBU by State") + theme(axis.text.x = element_text(angle = 90)) + geom_text(aes(label=MedianIBU), hjust=-.4, size=2.2) +   coord_flip()
```


## 5a State With Maximum ABV

```{r  echo=TRUE}
state <- brewandbeers[which.max(brewandbeers$ABV),][4]

```
`r state` has the highest ABV beer.

## 5b State With Maximum IBU
The following State has the maximum IBU beer.
```{r  echo=TRUE}
state <- brewandbeers[which.max(brewandbeers$IBU),][4]

```
`r state` has the highest IBU beer.

## 6 Summary Statistics and distribution of ABV.
```{r  echo=TRUE}
#This gets the quantile
#Display quantile as Table
abvquantile <- quantile(brewandbeers$ABV)
kable(abvquantile) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"), full_width = F, position="left")

#Draw a boxplot to explain the summary statistics

beers %>% ggplot(aes(y=ABV)) + geom_boxplot(fill="orange") + ggtitle("Summary Statistics of ABV Beer")
```

From the summary statistics, we find most of the beer ABV is between 5% and 6.7%. Interestingly there is one very low ABV beer (may be for recovering alcoholics) and there are a handful in higher ABV, though vast majority of people prefer betwen 5 and 7%.

## 7 ABV v IBU
Are there any relationship between ABV and IBU? Let us find from the following data:
```{r  echo=TRUE}
brewandbeers %>% filter(!is.na(IBU) & !is.na(ABV)) %>% ggplot(aes(x=ABV, y=IBU, color=IBU)) + geom_point()
```

While most of the beers are between 5% and 6.7% of ABV, and the IBU is under 50. There is some market on higher alcohol and higher IBU of 6% to 10% of ABV and 50 an 100 IBU. There is an linear relationship between ABV and IBU, which suggests people prefer a more balanced beer, where ABV and IBU are relatable. There are few beer that doesn't follow the trend. 
It suggests highest ABV and IBU are still produced, but the market is niche, and also the relationship stays even in very higher ABV vs IBU.



